HEP Software Foundation Community White Paper Working Group -- Training, Staffing and Careers

### 2018-06-13 CWP/papers/HSF-CWP-2017-02_training/latex/cwp-wg-training-careers.tex
Tutorial of the month.

An important point to consider is the difficulty, often experienced in the past,
in developing large software programs across different experiment communities.
While there already are experiment-specific training efforts in place, there are
many needs that are in common between experiments. Therefore, we should probably
strive to extract the common knowledge and build a common training programme. This
would reduce duplication of efforts and enable growth of a shared training
culture by accumulating and sharing expertise across different experiments.

To realise this goal, a possibility could be the institution of a {\em
federation} of existing training initiatives, a very lightwheight umbrella just
aimed at improving the efficiency and cost-effectivness of this important
activity. An appropriate incentive programme to reward those who train the community 
might help to facilitate the goals of such an initiative.

@@

The adoption of such an approach is made rather easy in WikiToLearn by the
relative simplicity of the wikimedia-based toolset: users contribute their
training material using just a web-browser, and in order to do this efficiently,
the necessary learning curve has been kept appropriately shallow.

An interesting exercise in this context has recently been made by colleagues of
the GridKa School of computing: the material from this year (2017) has been made
publicly available on WTL~\cite{WikiToLearnGridka}.
This constitutes an intersting example of what can be accomplished using this
platform; it is just a first example of what's possible, but an inspiring one.

>>>>

An important point to consider is the difficulty, often experienced in the past,
in developing large software programs across different experiment communities.
While there already are experiment-specific training efforts in place, there are
many needs that are in common between experiments. Therefore, we should probably
strive to extract the common knowledge and build a common training programme. This
would reduce duplication of efforts and enable growth of a shared training
culture by accumulating and sharing expertise across different experiments.
To realise this goal, a possibility could be the institution of a {\em
federation} of existing training initiatives, a very lightwheight umbrella just
aimed at improving the efficiency and cost-effectivness of this important
activity. An appropriate incentive programme to reward those who train the community 
might help to facilitate the goals of such an initiative.

When trying to exchange training materials within a group the first problem is
to convince authors to contribute their work. This issue
can be alleviated with cautious intellectual property clauses. Another
problem is that trainers do not like to reuse given material as-is. They usually
want to refactor it, building their own training story.
This makes it difficult to have everyone agree on common material.
It is even challenging to agree to host material in the same
centralized place. A way to overcome this could be to settle for a centralized
catalog, with each author being free to host their material in their location
of choice.

Another challenge is the wide heterogeneity in
student competence. Special care must be given to setup
a training structure that can cope with both beginners and advanced students.
This may rest on a low granularity subdivision of independant topics.
The IN2P3 authors try to restrict their tutorials to 20 minutes
of material, as inspired by the Pomodoro technique, so that one can easily jump
and compose one's own ``menu''. However, it turns out to be tricky to keep such
small tutorials really independant and meaningful when they can be selected at
will.

All trainers have also faced the tremendous loss of time
in software installation for any practical exercises as many students have
not managed to prepare their machines in advance.
Here containers may bring real
progress, provided that everyone has at least Docker installed on their machine.
JupyterHub, also, is a technology which enable training session with
no need for installation (but with a need for network).

@@

The adoption of such an approach is made rather easy in WikiToLearn by the
relative simplicity of the wikimedia-based toolset: users contribute their
training material using just a web-browser, and in order to do this efficiently,
the necessary learning curve has been kept appropriately shallow.
An interesting exercise in this context has recently been made by colleagues of
the GridKa School of computing: the material from this year (2017) has been made
publicly available on WTL~\cite{WikiToLearnGridka}.
This constitutes an interesting example of what can be accomplished using this
platform; it is just a first example of what's possible, but an inspiring one.

Another example of web-based platform is a collection of online tutorials
(mostly written in French) hosted on the Gitlab IN2P3
server~\cite{Code-Swim-Coaches}. Taking advantage of
the Gitlab ability to host Docker images, those tutorials aim to avoid the
``damned installation step'' that often absorbs half of a training session.
Similarly to the GridKa School for WikiToLearn, the annual IN2P3 Computing
Days are an opportunity to refresh and extend the collection of
tutorials every year. Future work will focus on English translation, and the
development of a web site which will index the above tutorials, together with
the best recommended external tutorials.
